Traffic agencies must freeze sparse sensor siting plans as long-lived fixed-infrastructure transportation network-design decisions before most future traffic states are observed, then reconstruct the hidden network state from partial measurements. We formulate that deployment-time question as a recoverability-driven bilevel stochastic network design problem rather than as coverage maximization or candidate-pool selection. The proposed method, TRACE-BiOpt, is a single recoverability-driven bilevel stochastic transportation network-design method centered on a transparent inverse problem: it chooses a fixed sensor set by minimizing one recoverability-driven objective whose lower level is a transparent GLS/MAP reconstruction problem and whose upper level combines hidden-state Huber reconstruction loss, posterior uncertainty, tail-risk, and spatial redundancy regularization. A deterministic solver uses scale-adaptive initialization and one-exchange refinement under the same objective; pre-registered baselines are never used as method candidates. Across PeMS7\_228, PeMS7\_1026, and Seattle, at 10\%, 20\%, and 30\% budgets over ten split seeds, TRACE-BiOpt obtains the lowest mean held-out GLS/MAP MAE against all 21 pre-registered non-BiOpt baselines spanning 11 method families in every tested dataset-budget regime. All nine rows now show submission-ready paired dominance, Holm-corrected one-sided paired tests across all 189 current-best row-baseline comparisons leave no statistically tied or significantly better challenger, and a bounded exact benchmark exact-hits 27/27 deterministic 16-node subnetworks under the same objective. The intended reading is that TRACE-BiOpt should be evaluated as one fixed-infrastructure bilevel stochastic transportation network-design method built around a transparent inverse problem and judged under an external audited comparison-class contract that compresses to row-wise strongest-baseline dominance on held-out reconstruction error and then survives all-baseline corrected paired tests. Global optimality and untested-baseline dominance remain explicitly excluded. The intended first reading is simple: one long-lived infrastructure decision, one transparent inverse problem, one deterministic solver, one scoped theory package, and one external audited comparison-class contract over 21 pre-registered baselines spanning 11 method families with no surviving challenger after 189 Holm-corrected paired tests.
