\section{Introduction}
\label{sec:introduction}

Traffic agencies must freeze sparse sensor siting plans before most future traffic states are observed, yet the operational question usually concerns the hidden full-network state that must later be inferred from partial measurements. Fixed loop detectors, cameras, probe feeds, and connected-vehicle observations therefore create a fixed-infrastructure transportation network design problem rather than a pure coverage problem. A sensor can be valuable because it makes unobserved links recoverable under a transparent estimation model, even when it is not the highest-flow or highest-variance location. In this paper, the design object is recoverability of the hidden network state rather than visibility of the observed subset.

This paper studies sparse traffic sensor siting as recoverability-driven stochastic network design for transparent state reconstruction. Given a fixed budget, we choose a fixed sparse set of nodes or links to instrument and then evaluate how well a transparent estimator reconstructs the hidden complement under future traffic uncertainty. This differs from classical count-location and observability formulations that emphasize origin-destination coverage, flow interception, link independence, partial/full link-flow observability, or sensor reliability \citep{yang1998optimal,gentili2011survey,hu2009identification,he2013graphical,viti2014partial,xu2016robust,salari2019failure}; it also differs from traffic state estimation and forecasting work that improves estimators while taking the observation pattern as fixed \citep{coric2012traffic,seo2017survey,li2018dcrnn,yu2018stgcn}. The siting plan and the reconstruction model must therefore be evaluated together as one transportation network design argument rather than as separate sensing and estimation steps.

We propose TRACE-BiOpt, a transparent recoverability-driven bilevel stochastic optimization method for fixed-infrastructure traffic sensor siting. The lower level is a GLS/MAP reconstruction problem that estimates the full traffic state from the chosen sensors. The upper level chooses the sensor set by minimizing one objective: hidden-state Huber reconstruction loss plus posterior uncertainty, tail-risk, and spatial redundancy terms. This design is deliberately not a portfolio or AutoML-style selector. Classical and reviewer-facing layouts such as random, top variance, degree, coverage, graph sampling, QR/POD, greedy A-trace, D-logdet, RCSS, validation-swap TRACE-SL, and multistart swap appear only as pre-registered baselines.

The intended reading is deliberately narrow: TRACE-BiOpt should be evaluated
as one fixed-infrastructure transportation network-design method
built around a transparent inverse problem and judged by whether it
achieves row-wise strongest-baseline dominance on held-out reconstruction error
and survives all-baseline corrected paired tests, not as a sparse-sensor
benchmark entry with a different winner on each table.
Operationally, this means one auditable design argument tested
against 21 pre-registered baselines spanning 11 method families, with no
surviving challenger after 189 Holm-corrected paired comparisons.

The contributions are fivefold.
First, we formulate sparse traffic sensing as a recoverability-driven bilevel stochastic network design problem rather than as coverage maximization, candidate-pool selection, or an estimator benchmark on a fixed graph.
Second, we define the lower level as a transparent GLS/MAP inverse problem and the upper level as one recoverability-driven objective over hidden-state loss, posterior uncertainty, tail-risk, and spatial redundancy.
Third, we develop TRACE-BiOpt as a deterministic single-objective solver with scale-adaptive initialization and one-exchange refinement under that same recoverability-driven objective.
Fourth, we give a theory package for the method: MAP closed form and stability, posterior trace as Bayes squared reconstruction risk, a formal CVaR tail-risk interpretation, uniform validation generalization over all size-$k$ layouts, and a computable exchange-stationarity certificate for the deterministic design solver.
Fifth, we provide a ten-seed experimental evaluation across PeMS7\_228, PeMS7\_1026, and Seattle at 10\%, 20\%, and 30\% budgets. TRACE-BiOpt achieves the lowest mean GLS/MAP MAE against the best pre-registered non-BiOpt baseline in all nine dataset-budget regimes; Holm-corrected one-sided paired tests across all 189 comparisons leave no statistically tied or significantly better challenger, and a bounded exact benchmark shows 27/27 exact hits on deterministic 16-node induced subnetworks under row-wise parameters. All claims are explicitly scoped to the tested baselines and regimes.

The contribution stack that this paper asks a TR-B reader to evaluate is as follows: the formulation is a bilevel recoverability-driven design problem, the algorithm is a single-objective deterministic solver, the theory package is scoped but nontrivial, and the empirical results are row-wise strongest-baseline, split-paired, and bounded-exact on deterministic subnetworks.

% Contribution stack table moved to supplementary material (SUBMISSION_EVIDENCE_INDEX.md)

The same story can be stated from a reviewer's perspective by answering the four questions that most directly separate a TR-B network-design contribution from a layout benchmark: is TRACE-BiOpt one method or a selector, does it beat the row-wise strongest pre-registered baseline, how far does the solver guarantee go, and what remains explicitly bounded.

% Reader guide table moved to supplementary material (SUBMISSION_EVIDENCE_INDEX.md)

\input{tables/table_trace_biopt_dominance}

The rest of the paper is organized as follows. Section~\ref{sec:related} positions TRACE-BiOpt relative to traffic count location, sensor selection, graph sampling, and traffic estimation. Section~\ref{sec:problem} defines the recoverability-driven bilevel stochastic network design problem. Section~\ref{sec:method} presents TRACE-BiOpt and the scoped theoretical statements. Section~\ref{sec:experiments} reports the main dominance results. Section~\ref{sec:mechanism} develops the mechanism and calibration diagnostics, Section~\ref{sec:robustness} separates bounded deployment-stress evidence from the main claim lane, Section~\ref{sec:discussion} discusses limitations, and Section~\ref{sec:conclusion} concludes.
