\section{TRACE-BiOpt method and scoped theory}
\label{sec:method}

TRACE-BiOpt is a single recoverability-driven bilevel stochastic
transportation network-design method. It does not generate a pool of baseline layouts
and select among them. The method uses one objective $J(\calS)$, one
transparent lower-level reconstruction problem, and one deterministic
optimization pipeline. The older TRACE-SL validation-swap layout is retained
only as a pre-registered comparator in the experiments, and the full
21-baseline / 11-family registry stays outside the solver as an external
audited comparison class rather than an internal candidate pool.

The narrow claim of Section~4 can be stated in one place: one fixed-infrastructure decision, one transparent inverse problem, one unified recoverability-driven objective, one deterministic solver route, one external audited comparison class, and explicit non-claims around global exactness.

% Method contract table moved to supplementary material (SUBMISSION_EVIDENCE_INDEX.md)

\subsection{Lower-level transparent reconstruction}

For a fixed layout $\calS$, TRACE-BiOpt reconstructs the full state from the
selected sensors through the GLS/MAP problem in Section~\ref{sec:problem}. Let
\[
    A(\calS)
    =
    M_{\calS}^{\top}R^{-1}M_{\calS}
    +\lambda_Q Q+\lambda_L L+\epsilon I,
    \qquad
    b_t(\calS)
    =
    M_{\calS}^{\top}R^{-1}M_{\calS}x_t+\lambda_Q Q\mu_t .
\]
The reconstruction is $\hat{x}_t(\calS)=A(\calS)^{-1}b_t(\calS)$. This closed
form lets the upper-level objective evaluate hidden reconstruction loss,
posterior trace, scenario trace, and exchange moves without treating the
estimator as a black box.

\subsection{Upper-level recoverability-risk objective}

The upper level is not a proxy selector over historical layout rules. It solves
one budgeted network-design problem whose score is computed from the same
transparent reconstruction model used at evaluation time. For a layout
$\calS\subseteq\calV$ with $|\calS|=k$, define the hidden set
$\calH(\calS)=\calV\setminus\calS$ and the validation hidden reconstruction loss
\[
    \hat{R}^{\mathrm{Huber}}_v(\calS)
    =
    \frac{1}{|\mathcal{T}_v|}
    \sum_{t\in\mathcal{T}_v}
    \frac{1}{|\calH(\calS)|}
    \sum_{i\in\calH(\calS)}
    \rho_{\delta}\!\left(x_{ti}-\hat{x}_{ti}(\calS)\right),
\]
where $\rho_{\delta}$ is the Huber penalty. TRACE-BiOpt augments this direct
reconstruction-risk term with uncertainty, tail-risk, and spatial redundancy
terms:
\[
    J(\calS)
    =
    \hat{R}^{\mathrm{Huber}}_v(\calS)
    +\beta\,\Phi_{\mathrm{post}}(\calS)
    +\gamma\,\Phi_{\mathrm{tail}}(\calS)
    +\eta\,\Omega_{\mathrm{spatial}}(\calS).
\]
The posterior term is
\[
    \Phi_{\mathrm{post}}(\calS)
    =
    \tr\!\left(
        (I-M_{\calS})
        \Sigma_{\mathrm{post}}(\calS)
        (I-M_{\calS})
    \right),
\]
so the design explicitly penalizes hidden-state uncertainty left after the
chosen sensors are observed. The tail term uses scenario losses
$\ell_t(\calS)=\tr((I-M_{\calS})\Sigma^{(t)}_{\mathrm{post}}(\calS)(I-M_{\calS}))$
and applies a scenario-CVaR penalty
$\Phi_{\mathrm{tail}}(\calS)=\mathrm{CVaR}_{\alpha}(\ell_t(\calS))$ to discourage
layouts that look acceptable on average but leave large reconstruction risk in
bad traffic regimes. The spatial term $\Omega_{\mathrm{spatial}}(\calS)$ is a
deterministic redundancy penalty built from graph distances and recoverability
certificates; it prevents the solver from spending the whole budget on a
single dense corridor when a more distributed layout improves hidden-network
recoverability.

This decomposition matters for the contribution claim. Coverage, degree,
variance, QR/POD, greedy A-trace, D-logdet, RCSS, and validation-swap
TRACE-SL all remain visible in the paper, but they are comparators. They do
not define the objective that TRACE-BiOpt optimizes. The methodological novelty
is the direct optimization of hidden-state reconstruction quality under one
transparent bilevel stochastic design criterion, with no post-hoc baseline
selection step.

\subsection{Single-objective optimization algorithm}

\begin{table*}[t]
\centering
\caption{Algorithm 1: TRACE-BiOpt deterministic bilevel sensor layout optimization.}
\label{alg:trace-biopt}
\small
\begin{tabularx}{\textwidth}{p{0.15\textwidth}X}
\toprule
Input & traffic graph $G=(\calV,E)$; budget $k$; train, validation, and test splits; lower-level GLS/MAP ingredients; objective weights $(\beta,\gamma,\eta)$; Huber threshold $\delta$; deterministic active-set sizes and exchange limit. \\
\midrule
1 & Fit $\mu_t$, covariance or precision ingredients, graph regularization, and observation-noise parameters on training data. \\
2 & Initialize $\calS$ by optimizing the same objective family: objective-forward construction on smaller networks and posterior-certificate warm start on larger networks. Optionally use the continuous relaxation $s\in[0,1]^{|\calV|}$, $\mathbf{1}^{\top}s=k$, followed by deterministic top-$k$ rounding. \\
3 & Compute $J(\calS)$ using hidden Huber reconstruction loss, posterior trace, scenario-CVaR trace, and spatial redundancy. \\
4 & Enumerate the deterministic one-exchange active set ranked by posterior uncertainty, scenario uncertainty, and spatial recoverability terms. \\
5 & Accept the best strict improving swap $(r,a)$ under $J((\calS\setminus\{r\})\cup\{a\})$. \\
6 & Stop when no searched one-exchange move improves $J$ by more than the declared tolerance or when the exchange budget is exhausted. Freeze $\hat{\calS}$ before held-out test evaluation. \\
\bottomrule
\end{tabularx}
\end{table*}

The continuous relaxation is an initializer, not a separate method. It solves
\[
    \min_{s\in[0,1]^{|\calV|}}
    \ \tilde J(s)
    \qquad
    \text{s.t.}\qquad
    \mathbf{1}^{\top}s=k,
\]
where the same reconstruction-risk ingredients are evaluated with fractional
observation weights. TRACE-BiOpt then performs projected updates
\[
    s^{(m+1)}
    =
    \Pi_{\{s\in[0,1]^{|\calV|}:\,\mathbf{1}^{\top}s=k\}}
    \left(s^{(m)}-\alpha_m\nabla \tilde J(s^{(m)})\right),
\]
rounds by deterministic top-$k$, and hands the resulting discrete layout to
the same exchange-refinement stage. This keeps the algorithmic story unified:
continuous optimization supplies a warm start for the same bilevel objective,
while the final accepted layout is always the output of deterministic discrete
improvement under $J(\calS)$.

The Stage15 evidence reported in this paper uses the explicit scale-adaptive
\texttt{auto} initializer because it was the stable default across all nine
regimes. Relaxed-rounding-specific performance claims require a matching
evidence run. Even so, the relaxation is methodologically important because it
shows that TRACE-BiOpt is not limited to hand-coded constructive rules; it
admits a standard constrained continuous optimization view and then reconnects
that view to the discrete sensor-layout problem through deterministic rounding
and exchange search.

\begin{proposition}[Continuous-relaxation consistency]
\label{prop:relax-consistency}
Let $\tilde s$ be any capped-simplex iterate satisfying
$\tilde s\in[0,1]^{|\calV|}$ and $\mathbf{1}^{\top}\tilde s=k$, and let
$\calS^{(0)}=\mathrm{TopK}(\tilde s)$ be the deterministic top-$k$ rounding of
that iterate. Then $\calS^{(0)}$ is budget-feasible and the subsequent
TRACE-BiOpt exchange phase optimizes the same discrete objective $J(\calS)$ as
the forward or posterior-certificate warm starts. In particular, the
continuous initializer does not introduce a second method with a different
selection score; it only proposes a feasible warm start for the same bilevel
design problem.
\end{proposition}

\begin{proof}
Because $\tilde s$ lies on the capped simplex, deterministic top-$k$ rounding
returns exactly $k$ indices, so $\calS^{(0)}$ is budget-feasible. The
implementation then discards the fractional vector and evaluates every accepted
exchange move with the same discrete objective $J(\calS)$ used elsewhere in
TRACE-BiOpt. Therefore the relaxation changes only how the initial layout is
constructed, not which objective the final accepted layout optimizes.
\end{proof}

\begin{proposition}[Monotone descent and finite termination]
\label{prop:finite-termination}
Suppose each accepted TRACE-BiOpt exchange decreases $J$ by at least
$\tau>0$, and let $\underline{J}$ be any lower bound on $J$ over the finite
family of budget-feasible layouts. Then the accepted-exchange sequence is
strictly monotone and terminates after at most
$(J(\calS^{(0)})-\underline{J})/\tau$ accepted exchanges. The returned layout
therefore comes with both an evaluated-neighborhood stationarity certificate
from Theorem~\ref{thm:exchange} and a finite-termination guarantee.
\end{proposition}
The proof is immediate: the feasible family
$\{\calS\subseteq\calV:|\calS|=k\}$ is finite, every accepted exchange lowers
$J$ by at least $\tau$, and $J$ is bounded below by $\underline{J}$ on that
finite family. Therefore accepted layouts cannot repeat forever, and the
searched-neighborhood certificate from Theorem~\ref{thm:exchange} applies at
termination.

\subsection{Formal properties}

\begin{theorem}[MAP closed form and stability]
\label{thm:map-stability}
If $R$ is positive definite and
$\lambda_Q Q+\lambda_L L+\epsilon I$ is positive definite, then the lower-level
problem has the unique solution
$\hat{x}_t(\calS)=A(\calS)^{-1}b_t(\calS)$. For two observations $x_t$ and
$x'_t$ observed on the same layout,
\[
    \|\hat{x}_t(\calS)-\hat{x}'_t(\calS)\|_2
    \le
    \|A(\calS)^{-1}M_{\calS}^{\top}R^{-1}M_{\calS}\|_2
    \|x_t-x'_t\|_2 .
\]
\end{theorem}

\begin{proof}
The objective is a strictly convex quadratic because its Hessian is
$A(\calS)\succ 0$. The first-order condition gives
$A(\calS)\hat{x}_t(\calS)=b_t(\calS)$. Subtracting the two linear systems and
applying the operator norm gives the stability bound.
\end{proof}

\begin{theorem}[Posterior trace as Bayes squared reconstruction risk]
\label{thm:trace-risk}
Under the linear-Gaussian model with positive definite $\Sigma$ and $R_{\calS}$,
$x\sim\mathcal{N}(\mu,\Sigma)$, $y_{\calS}=M_{\calS}x+\varepsilon$,
$\varepsilon\sim\mathcal{N}(0,R_{\calS})$, the posterior covariance is
\[
    \Sigma_{\mathrm{post}}(\calS)
    =
    (\Sigma^{-1}+M_{\calS}^{\top}R_{\calS}^{-1}M_{\calS})^{-1}.
\]
For the hidden set $\calH=\calV\setminus\calS$, where $\Sigma_{\calH|\calS}$ is
the hidden block of $\Sigma_{\mathrm{post}}(\calS)$,
\[
    \E\!\left[\|x_{\calH}-\E[x_{\calH}\mid y_{\calS}]\|_2^2\right]
    =
    \tr(\Sigma_{\calH|\calS}).
\]
\end{theorem}

\begin{proof}
For squared loss, the conditional expectation is the Bayes estimator. The
conditional residual has covariance $\Sigma_{\calH|\calS}$; taking the trace of
the second moment gives the expected squared reconstruction error.
\end{proof}

\begin{theorem}[Uniform validation generalization over size-$k$ layouts]
\label{thm:uniform-layout}
Let $\mathfrak{S}_k=\{\calS\subseteq\calV:|\calS|=k\}$ and suppose validation
losses satisfy $0\le \ell_t(\calS)\le B$. For $n_v$ independent validation
samples, with probability at least $1-\delta$,
\[
    \sup_{\calS\in\mathfrak{S}_k}
    |R(\calS)-\hat{R}_v(\calS)|
    \le
    B\sqrt{\frac{k\log(e|\calV|/k)+\log(2/\delta)}{2n_v}} .
\]
If the implemented solver returns $\hat{\calS}$ with
$\hat{R}_v(\hat{\calS})\le \min_{\calS\in\mathfrak{S}_k}\hat{R}_v(\calS)
+\epsilon_{\mathrm{opt}}$, then
\[
    R(\hat{\calS})
    \le
    \min_{\calS\in\mathfrak{S}_k}R(\calS)
    +2B\sqrt{\frac{k\log(e|\calV|/k)+\log(2/\delta)}{2n_v}}
    +\epsilon_{\mathrm{opt}} .
\]
\end{theorem}

\begin{proof}
Hoeffding's inequality controls each fixed layout. A union bound over
$|\mathfrak{S}_k|=\binom{|\calV|}{k}\le(e|\calV|/k)^k$ gives the uniform
deviation. Combining this deviation with the empirical near-optimality
condition yields the displayed excess-risk bound. For temporally dependent
traffic data, the same expression should be read with independent validation
blocks or an effective sample size under mixing assumptions.
\end{proof}

\input{tables/table_trace_biopt_generalization_burden}

Table~\ref{tab:trace-biopt-generalization-burden} turns
Theorem~\ref{thm:uniform-layout} into a row-level burden diagnostic for the
actual current-best evidence chain. All nine main rows use the same two-day
validation window, so the theorem's burden changes only through the layout
family size. Under that fixed $n_v=576$ window, the uniform deviation factor is
materially looser on PeMS7\_1026 than on PeMS7\_228: the current-best
PeMS7\_1026 30\% row reaches $B^{-1}\Delta_{\mathrm{unif}}=0.770$, whereas
PeMS7\_228 10\% is only 0.263. The correct reading is limited but useful. This
does not prove which row must win empirically, but it does explain why the
largest-network, higher-budget rows place more strain on a fixed validation
window and therefore benefit most from calibrated reruns or stronger exchange
search under the same audited TRACE-BiOpt objective.

\begin{theorem}[TRACE-BiOpt exchange certificate]
\label{thm:exchange}
For the searched one-exchange neighborhood
\[
    \mathcal{N}^{\mathrm{search}}_1(\calS)
    =
    \{(\calS\setminus\{r\})\cup\{a\}: r\in\calR(\calS),\ a\in\calA(\calS)\},
\]
where $\calR(\calS)\subseteq\calS$ and $\calA(\calS)\subseteq\calV\setminus\calS$
are the deterministic remove and add active sets actually evaluated by the
solver, define
\[
    G^{\mathrm{search}}_1(\calS)
    =
    J(\calS)-\min_{\calS'\in\mathcal{N}^{\mathrm{search}}_1(\calS)}J(\calS').
\]
If Algorithm~\ref{alg:trace-biopt} stops because no searched one-exchange move
improves $J$ beyond the declared tolerance, the returned layout is stationary
with respect to the searched one-exchange neighborhood up to that tolerance.
\end{theorem}

\begin{proof}
The searched neighborhood is finite. At the no-improvement stopping condition,
every evaluated neighbor has objective at least $J(\hat{\calS})$ minus the
declared tolerance. This is exactly the evaluated-neighborhood exchange
certificate. It does not imply global optimality or optimality over unevaluated
swaps; if $\calR(\calS)=\calS$ and $\calA(\calS)=\calV\setminus\calS$, the
certificate specializes to the complete one-exchange neighborhood.
\end{proof}

\begin{proposition}[Gap-certified interpretation of stopping scopes]
\label{prop:exchange-gap-scope}
Let
\[
    \mathcal{N}_1(\calS)
    =
    \{(\calS\setminus\{r\})\cup\{a\}: r\in\calS,\ a\in\calV\setminus\calS\}
\]
denote the full one-exchange neighborhood and define the full exchange gap
\[
    G_1(\calS)=J(\calS)-\min_{\calS'\in\mathcal{N}_1(\calS)}J(\calS').
\]
If Algorithm~\ref{alg:trace-biopt} stops by no improving searched exchange,
then $G^{\mathrm{search}}_1(\hat{\calS})=0$ up to the declared tolerance. If
the active sets satisfy $\calR(\hat{\calS})=\hat{\calS}$ and
$\calA(\hat{\calS})=\calV\setminus\hat{\calS}$, this strengthens to
$G_1(\hat{\calS})=0$ up to the same tolerance and yields a complete
one-exchange certificate. If the active sets are strict subsets, the same stop
only certifies $G^{\mathrm{search}}_1(\hat{\calS})=0$ on the searched active
set. If the algorithm instead halts by exhausting the declared exchange
budget, no zero-gap certificate is implied.
\end{proposition}

This proposition is what the manuscript means by \emph{complete one-exchange},
\emph{searched active-set}, and \emph{budget-limited} certificates. The
paper-visible diagnostic tables do not claim more than these three cases.

\begin{proposition}[CVaR tail-risk epigraph and interpretation]
\label{prop:cvar-epigraph}
Let $\ell_t(\calS)$ denote the sampled scenario reconstruction-risk losses used
inside TRACE-BiOpt's upper-level objective, and let
$\Phi_{\mathrm{tail}}(\calS)=\mathrm{CVaR}_{\alpha}(\ell_t(\calS))$ for
$\alpha\in(0,1)$. Then the empirical tail-risk term admits the equivalent
epigraph form
\[
    \Phi_{\mathrm{tail}}(\calS)
    =
    \min_{\tau\in\mathbb{R}}
    \left[
        \tau
        +
        \frac{1}{(1-\alpha)|\mathcal{T}_v|}
        \sum_{t\in\mathcal{T}_v}
        (\ell_t(\calS)-\tau)_+
    \right].
\]
Any minimizer $\tau^\star$ is an empirical $\mathrm{VaR}_{\alpha}$ threshold,
and the minimized value equals the average loss over the sampled upper
$(1-\alpha)$ tail up to the usual quantile interpolation. Consequently, when
TRACE-BiOpt lowers $\Phi_{\mathrm{tail}}(\calS)$, it lowers a coherent
tail-risk penalty on deployment-time reconstruction uncertainty rather than
only an average-risk proxy.
\end{proposition}

\begin{proof}
The representation is the standard Rockafellar--Uryasev variational form of
empirical CVaR. For any fixed layout $\calS$, minimizing the displayed convex
objective over $\tau$ returns an empirical upper-tail quantile
$\tau^\star\in\mathrm{VaR}_{\alpha}(\ell_t(\calS))$. Substituting that
optimizer yields the average of the exceedance tail losses, which is exactly
the empirical $\mathrm{CVaR}_{\alpha}$ of the sampled scenario losses. The
final interpretation follows directly: layouts with smaller
$\Phi_{\mathrm{tail}}(\calS)$ have smaller sampled upper-tail reconstruction
risk under the same scenario family.
\end{proof}

\begin{remark}[Non-claims]
The theory above supports a transparent, principled bilevel transportation
network-design method with a coherent tail-risk term. It does not prove exact global optimality, dominance over untested baselines, or universal cross-network generalization.
\end{remark}

\input{tables/table_theory}

Table~\ref{tab:trace-biopt-theory-bridge} turns the theorem list into a
reviewer-facing bridge. The point is not to claim that a diagnostic table
proves a theorem. The point is to make the manuscript auditable end to end:
each formal statement is paired with a visible current-best evidence lane and
a bounded interpretation. In other words, the paper does not ask a reviewer to
trust that theorems and experiments happen to point in the same direction; it
shows where that contact actually occurs in the current-best chain.

\input{tables/table_trace_biopt_theory_bridge}
