\section{Ablations and robustness routing}
\label{sec:analysis}

The ablation contract separates the roles of random candidate pools, validation selection, certificate-guided candidate generation, RCSS selection, validation-aware local refinement, and multistart swap search. Table~\ref{tab:ablation} shows the PeMS7\_228 component comparison at 10\%, 20\%, and 30\% budgets. At 10\%, TRACE-SL improves over random mean, validation-selected random, and greedy A-trace, but the multistart swap comparator is statistically indistinguishable. At 20\% and 30\%, local refinement gives clearer gains over the reviewer-facing baselines and over the multistart row at 30\%.

\input{tables/table_ablation}

The ablation should be interpreted at the component level rather than as a kitchen-sink heuristic. Greedy A-trace and scenario risk variants are certificate-only candidates. Validation-selected random isolates validation selection without certificate-guided candidate generation. RCSS-selected layouts isolate score-based selection before local refinement. Validation-swap selected layouts add the local neighborhood search. Multistart swap is retained as a serious comparator because it is close at low budget and strong enough to prevent an overbroad dominance claim.

Robustness evidence is routed as stress-test evidence. The project includes perturbation artifacts for sensor failure, observation noise, missingness, heterogeneous costs, temporal shift, and candidate-pool sensitivity. These tests support bounded robustness discussion in the tested settings. They do not support the phrase globally robust, and they do not replace held-out Stage12 evidence for main performance claims.
