\section{Conclusion}
\label{sec:conclusion}

TRACE-BiOpt reframes sparse traffic sensor deployment as a recoverability-driven bilevel network design problem. Instead of ranking locations by coverage, variance, or topology alone, it selects a fixed sensor set by directly optimizing hidden-network reconstruction risk under a transparent GLS/MAP estimator.

That is also the comparison class the paper asks a reviewer to use: fixed-infrastructure transportation network design under downstream recoverability, not estimator tuning, benchmark-style model ranking, or post-hoc layout selection.

The method combines hidden Huber reconstruction loss, posterior uncertainty, scenario tail risk, and spatial regularization in one objective, then applies deterministic initialization and exchange refinement under that objective. Across PeMS7\_228, PeMS7\_1026, and Seattle current-best evidence, TRACE-BiOpt beats the best pre-registered non-BiOpt baseline in mean held-out GLS/MAP MAE in every tested dataset-budget regime, the audited comparison class spans 21 baselines across 11 method families, the current nine-row chain is submission-ready under paired audits, Holm-corrected one-sided paired tests across all 189 current-best row-baseline comparisons leave no statistically tied or significantly better pre-registered challenger, and the bounded exact benchmark exact-hits 27/27 deterministic 16-node subnetworks with zero objective gap. The final contribution is an auditable bridge between traffic infrastructure design, transparent inverse problems, robust bilevel stochastic optimization, and empirical sensor deployment evidence.

Stated in the most compressed TR-B form, the paper claims one thing: one
fixed-infrastructure transportation network-design method built around a
transparent inverse problem can be made formulation-consistent, theory-backed,
compressible to row-wise strongest-baseline dominance, multiplicity-robust
against the full audited comparison class, and interpretable as a principled
bilevel stochastic design method on the tested traffic networks without
turning into a post-hoc selector story.

Table~\ref{tab:trace-biopt-planning-takeaways} turns that bridge into a
network-level planning readout. The point is not to add another dominance
table. The point is to close with what a transport-planning reader can
actually do with the evidence: treat PeMS7\_1026 as a search-budget-sensitive
large network, PeMS7\_228 as an expansion-friendly calibrated rollout, and
Seattle as a conservative staged-rollout case whose 20/30\% promoted routes
are deployable while the 10\% row remains intentionally fail-closed on
promotion evidence.

\input{tables/table_trace_biopt_planning_takeaways}

In that sense, the paper's claim is not just that one siting heuristic wins a
benchmark. It is that a recoverability-driven network-design formulation, a
single-objective deterministic solver, a scoped but explicit theory package,
and an external audited comparison-class current-best evidence chain can be
made to line up in one transport-network design argument without hiding behind
a candidate pool.
The intended reading is therefore that TRACE-BiOpt should be evaluated as one
fixed-infrastructure transportation network-design method built around a
transparent inverse problem and judged under an external audited
comparison-class contract that compresses to row-wise strongest-baseline
dominance on held-out reconstruction error and then survives all-baseline
corrected paired tests, not as a sparse-sensor benchmark paper that happens to
borrow traffic data.
Its shortest reviewer-facing summary is the same one used at the front of the
paper: one long-lived infrastructure decision, one transparent inverse
problem, one deterministic solver, one scoped theory package, and one
external audited comparison-class contract over 21 pre-registered baselines
spanning 11 method families with no surviving challenger after 189
Holm-corrected paired tests.
For a transportation agency, the practical message is equally narrow and
concrete: freeze the sparse sensor siting plan as an infrastructure design decision
before deployment, evaluate it by the recoverability it creates for the hidden
network state under a transparent estimator, and compare it to the strongest
audited alternative rather than to a weak fixed baseline or a post-hoc
estimator gain. That is the decision contract this paper argues for.
