\documentclass[a4paper,fleqn]{cas-sc}

\usepackage[authoryear,longnamesfirst]{natbib}
\usepackage{amsmath,amssymb,amsthm,booktabs,array,tabularx}
\usepackage{graphicx}
\usepackage{xurl}

\input{math_commands}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}

\shorttitle{TRACE-BiOpt Transportation Network Design}
\shortauthors{Anonymous Authors}

\title[mode=title]{TRACE-BiOpt: Recoverability-Driven Bilevel Transportation Network Design for Sparse Traffic Sensor Siting}

\author[1]{Anonymous Author}
\affiliation[1]{organization={Anonymous Institution},
            city={Anonymous City},
            country={Anonymous Country}}

\begin{abstract}
Traffic agencies must freeze sparse sensor siting plans as long-lived fixed-infrastructure transportation network-design decisions before most future traffic states are observed, then reconstruct the hidden network state from partial measurements. We formulate that deployment-time question as a recoverability-driven bilevel stochastic network design problem rather than as coverage maximization or candidate-pool selection. The proposed method, TRACE-BiOpt, is a single recoverability-driven bilevel stochastic transportation network-design method centered on a transparent inverse problem: it chooses a fixed sensor set by minimizing one recoverability-driven objective whose lower level is a transparent GLS/MAP reconstruction problem and whose upper level combines hidden-state Huber reconstruction loss, posterior uncertainty, tail-risk, and spatial redundancy regularization. A deterministic solver uses scale-adaptive initialization and one-exchange refinement under the same objective; pre-registered baselines are never used as method candidates. Across PeMS7\_228, PeMS7\_1026, and Seattle, at 10\%, 20\%, and 30\% budgets over ten split seeds, TRACE-BiOpt obtains the lowest mean held-out GLS/MAP MAE against all 21 pre-registered non-BiOpt baselines spanning 11 method families in every tested dataset-budget regime. All nine rows now show submission-ready paired dominance, Holm-corrected one-sided paired tests across all 189 current-best row-baseline comparisons leave no statistically tied or significantly better challenger, and a bounded exact benchmark exact-hits 27/27 deterministic 16-node subnetworks under the same objective. The paper therefore claims one unified, analyzable, theory-backed bilevel stochastic transportation network-design method whose evidence contract is an external audited comparison-class contract that compresses to row-wise strongest-baseline dominance on the tested datasets and then survives all-baseline corrected paired tests, with global optimality and untested-baseline dominance explicitly excluded. The intended first reading is simple: one long-lived infrastructure decision, one transparent inverse problem, one deterministic solver, one scoped theory package, and one external audited comparison-class contract over 21 pre-registered baselines spanning 11 method families with no surviving challenger after 189 Holm-corrected paired tests.
\end{abstract}

\begin{highlights}
\item We cast sparse traffic sensing as fixed-infrastructure transportation network design for hidden-state recoverability.
\item TRACE-BiOpt commits one fixed layout under one recoverability-driven bilevel stochastic objective with a transparent inverse problem.
\item Current-best paired audits show strongest-baseline dominance on all nine tested dataset-budget rows, and Holm-corrected paired tests leave no statistically tied or better challenger across 21 pre-registered baselines spanning 11 method families.
\item The theory package covers MAP stability, posterior risk identity, CVaR tail-risk interpretation, layout-class generalization, and exchange-stationarity certificates.
\item A bounded exact benchmark exact-hits 27/27 deterministic 16-node subnetworks under the same objective.
\end{highlights}

\begin{keywords}
transportation network design \sep traffic sensor siting \sep traffic state reconstruction \sep bilevel stochastic optimization \sep transparent inverse problem
\end{keywords}

% The CAS front matter emits template-level overfull/empty-anchor noise from
% its zero-width keyword/sidebar machinery; suppress only the maketitle call.
\newdimen\savedhfuzz
\savedhfuzz=\hfuzz
\hfuzz=120pt
\makeatletter
\let\savedHyWarning\Hy@Warning
\def\Hy@Warning#1{}
\makeatother
\maketitle
\makeatletter
\let\Hy@Warning\savedHyWarning
\makeatother
\hfuzz=\savedhfuzz

\input{sections/1_introduction}
\input{sections/2_related_work}
\input{sections/3_problem}
\input{sections/4_method_theory}
\input{sections/5_experiments}
\input{sections/6_ablation_robustness}
\input{sections/7_robustness}
\input{sections/7_discussion}
\input{sections/8_conclusion}

\section*{CRediT authorship contribution statement}
Anonymous for review.

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Data and code availability}
The manuscript reports curated aggregate artifacts and generation scripts under \path{TRC-23-02333/trace_sl_results/}. The primary TRACE-BiOpt result is generated from \path{current_best_trace_biopt_evidence/}, which now promotes all six PeMS dataset-budget rows and Seattle 20/30\% through complete replaceable Stage16 calibrated reruns, while Seattle 10\% intentionally remains on the Stage15 main evidence path. Raw traffic datasets are local inputs and are not distributed with the paper source; the source package records the expected local file layout, launch scripts, aggregate CSV/JSON contracts, and audit artifacts used to regenerate paper-visible tables. The reported numbers should therefore be read as reproducible from the curated artifacts, while full raw-data reproduction requires obtaining the underlying traffic datasets separately.

\bibliographystyle{cas-model2-names}
\bibliography{references}

\appendix
\input{sections/A_appendix}

\end{document}
